---
title: "GSS challange More Data"
author: "Leonel D. Bautista"
date: "9/28/23"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

$H_0$ There is no difference in the data when comparing race and attitude toward capital punishment.   

$H_A$ There is a difference in when comparing the race and attitude toward capital punishment.   

IF there is a difference in attitude, where in the data set is the difference?   

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

```{r}
head(mydata)
```

Looking at our data we see that the problem is Categorical~Categorical meaning that we will need to use the Cat_Cat helper file from the CandP set.   

We will expect to use a bar chart to visually express our data and a table to show the numerical summary of our data.   

Finally for our inferential results we will use a Chi-Squared test, this will only tell us if there is a difference and not where the difference is.
If there is a difference in the data, to identify its sourse we will run a FisherXY test, which loops the Fisher Test for all combinations.   

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")

```

We see in the graphical representation of our data that a disproportionate amount of white people are in favor of capital punishment.   

We do not see this trend when looking looking at other races, in fact more black people oppose capital punishment.   

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
```

As scene here, almost double the percentage of white people are in favor of capitol punishment.   

The black and other races are much closer to a 50:50 split.   

## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
```

In the above Chi-Squared test we see that we both have Chi-Squared Statistic much higher than the critical value given our degrees of freedom.  

This higher number shows us that most of the data set is different from its expected value.    

The p-value us also extremely small at $2.2 \text{ x } 10^{-16}$
This tells us that there is absolutely a difference somewhere in the data, to find where we will need a Fisher Test.   

```{r}
fisherXY(mydata$race,mydata$cappun,printAll = FALSE,confL = 0.99)
```

We found the difference, while running a Fisher Test at with a confidence level of 99% we see where the critical p-value is.   

The relationship between the races; white and black is where the greatest difference is with the previous small p-value of $2.2 \text{ x } 10^{-16}$   

We also wee that most of the p-values are lower than 0.05 but this has little application here as one of the variables is dramatically different.   

## More data version


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
